---
title: COVID-19 Radar 
author: by the TwinsUK data analysis Team
output:
  html_document:
      toc: true
      number_sections: true
      theme: readable
      highlight: textmate
      fig_width: 10
      fig_height: 3
      fig_caption: yes
      self_contained: true
---


# Data collected

As of `r day` the COVID-19 radar application got data for `r npatient` individuals, `r nfemale` of whom were females (`r pfemale`%) and `r nmale` males (`r pmale`%).
Their age ranged from `r minage` to `r maxage` year old, with a median of `r medianage` (mean=`r meanage`, SD=`r sdage`). `r sum(patient$contact_health_worker == "t")` contacted a health worker.


When we look at different age brackets, this is what we see:

````{r, echo=FALSE}
age.brackets
````

We also calculated their body mass index (or [BMI](https://en.wikipedia.org/wiki/Body_mass_index)) which had a median of `r medianbmi` kg/m^2 (mean=`r meanbmi`, SD=`r sdbmi`).


## Smoking status

We also asked whether they were smokers:


````{r, echo=FALSE}
diseases$is_smoker$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
diseases$is_smoker$brackets
````


## Disease status

We asked the app users whether they were suffering from a set of diseases, which are reported in the following.


### Lung diseases

These are our stats for all the individuals who participated:

````{r, echo=FALSE}
diseases$has_lung_disease$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
diseases$has_lung_disease$brackets
````


### Kidney diseases

These are our stats for all the individuals who participated:

````{r, echo=FALSE}
diseases$has_kidney_disease$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
diseases$has_kidney_disease$brackets
````

### Diabetes

These are our stats for all the individuals who participated:

````{r, echo=FALSE}
diseases$has_diabetes$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
diseases$has_diabetes$brackets
````

### Cancer

These are our stats for all the individuals who participated:

````{r, echo=FALSE}
diseases$has_cancer$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
diseases$has_cancer$brackets
````


## Medications

We asked the app users whether they were taking specific classes of medication, which are reported in the following.


### Corticosteroids

We asked whether they regularly took "NSAIDs" like ibuprofen, nurofen, diclofenac, naproxen, and these are our stats:

````{r, echo=FALSE}
medications$takes_corticosteroids$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
medications$takes_corticosteroids$brackets
````

### Blood pressure medications

**ACE inhibitor**

We asked whether they regularly took blood pressure medications ending in "-pril", such as enalapril, ramipril, perindopril, and these are our stats:

````{r, echo=FALSE}
medications$takes_blood_pressure_medications$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
medications$takes_blood_pressure_medications$brackets
````


**Angiotensin receptor blocker**

We asked whether they regularly took blood pressure medications ending in "-sartan", such as losartan, valsartan, irbesartan, and these are our stats:

````{r, echo=FALSE}
medications$takes_blood_pressure_medications_sartan$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
medications$takes_blood_pressure_medications_sartan$brackets
````


### Aspirin

We asked whether they regularly took aspirin (either baby aspirin or standard dose), and these are our stats:


````{r, echo=FALSE}
medications$takes_aspirin$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
medications$takes_aspirin$brackets
````


### Chemotherapy or immunotherapy for cancer

These are our stats for all the individuals who participated:

````{r, echo=FALSE}
medications$does_chemiotherapy$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
medications$does_chemiotherapy$brackets
````


### Immunosuppressant

We asked whether they regularly took immunosuppressant medications (including steroids, methotrexate, biologics), and these are our stats:


````{r, echo=FALSE}
medications$takes_immunosuppressants$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
medications$takes_immunosuppressants$brackets
````


# COVID-19 status

## Self-reported 

We asked the app users whether they believed they have already had COVID-19, and `r sum(!is.na(patient$already_had_covid))` answered. Please note, that is information is self-reported and is not confirmed by a positive test. 

`r nrow(already_had_covid)` reported thinking of have already had COVID-19, that is `r palready_had_covid`%.


When we look at different age brackets, this is what we see:

````{r, echo=FALSE}
age.brackets.already_had_covid
````


## Confirmed COVID-19 cases

`r nrow(positive)+nrow(negative)` were tested for COVID-19, and `r nrow(positive)` confirmed the disease (`r ppositive`%).
`r sum(positive$gender==0)` were females (`r pfemale.positive`%).

`r nrow(waiting)` individuals are still waiting for the results of the test.

When we look at different age brackets, this is what we see:

````{r, echo=FALSE}
age.brackets.positive
````


# What are the reported symptoms on `r day`?

`r sum(!is.na(assessment$health_status) & assessment$health_status == "healthy")` individuals told us that they were healthy, but `r sum(!is.na(assessment$health_status) & assessment$health_status == "not_healthy")` reported that they were not feeling quite right.

````{r, echo=FALSE}
health_status$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
health_status$brackets
````

We then asked the people that were not feeling fine which were their symptoms, and we details about them below:


## Fever

`r length(temperature)` individuals reported their temperature. The median value was `r median.temperature` (mean=`r mean.temperature`, SD=`r sd.temperature`). `r sum(temperature > 38)` had a temperature higher than 38 C.

`r sum(!is.na(assessment$fever) & assessment$fever == "t")` instead reported of having fever -- albeit not everyone of them reported their temperature:

````{r, echo=FALSE}
symptoms$fever$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$fever$brackets
````

## Persistent cough

````{r, echo=FALSE}
symptoms$persistent_cough$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$persistent_cough$brackets
````


## Shortness of breath

Out of `r tot.shortness_of_breath` who answered regarding shortness of breath, `r having.shortness_of_breath` reported experiencing it (`r rp(having.shortness_of_breath, tot.shortness_of_breath)`). For `r mild.shortness_of_breath` it was mild (`r rp(mild.shortness_of_breath, having.shortness_of_breath)`%), for `r significant.shortness_of_breath` it was significant (`r rp(significant.shortness_of_breath, having.shortness_of_breath)`%), and for `r severe.shortness_of_breath` it was severe (`r rp(severe.shortness_of_breath, having.shortness_of_breath)`%).

Taking into account as those reporting having shortness of breath, this is what we observe:

````{r, echo=FALSE}
shortness_of_breath$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
shortness_of_breath$brackets

````

## Chest pain


````{r, echo=FALSE}
symptoms$chest_pain$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$chest_pain$brackets
````


## Sore throat


````{r, echo=FALSE}
symptoms$sore_throat$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$sore_throat$brackets
````



## Hoarse voice


````{r, echo=FALSE}
symptoms$hoarse_voice$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$hoarse_voice$brackets
````


## Headache


````{r, echo=FALSE}
symptoms$headache$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$headache$brackets
````


## Loss of sense of smell


````{r, echo=FALSE}
symptoms$loss_of_smell$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$loss_of_smell$brackets
````

## Fatigue

Out of `r tot.fatigue ` who answered regarding experiencing fatigue, `r having.fatigue` reported experiencing it. For `r mild.fatigue` it was mild (`r rp(mild.fatigue, having.fatigue)`%), and for `r severe.fatigue` it was severe (`r rp(severe.fatigue, having.fatigue)`%).

Taking into account as those reporting having fatigue, this is what we observe:


````{r, echo=FALSE}
fatigue$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
fatigue$brackets
````


## Diarrhoea


````{r, echo=FALSE}
symptoms$diarrhoea$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$diarrhoea$brackets
````

## Abdominal pain

````{r, echo=FALSE}
symptoms$abdominal_pain$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$abdominal_pain$brackets
````

## Skipped meals

````{r, echo=FALSE}
symptoms$skipped_meals$general
````

and this is the breakdown for age:

````{r, echo=FALSE}
symptoms$skipped_meals$brackets
````


